---
title: "Homework 8"
author: "Sam Reeves"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. A company buys 100 lightbulbs, each of which has an exponential lifetime of 1000 hours.  What is the expected time for the first of these bulbs to burn out?

I accidentally chose this problem for the discussion board...  The expected value is "just" mu / n.

```{r}
n <- 100
mu <- 1000

mu / n
```


## 2. Assume that $X_1$ and $X_2$ are independent random variables, each having an exponential density with parameter $\lambda$.  Show that $Z = X_1 - X_2$ has density

## $f(z) = \frac{1}{2} \lambda e^{-\lambda |z|}$.

Because the two distributions have the same shape, the majority of values in $Z$ will always be distributed around zero.  As the lambda parameter of one $X_1$ approaches that of $X_2$, the distribution of $Z$ will approach a uniform normal, and it will always be centered at zero.

```{r}
n <- 100000
lambda <- 1
X_1 <- rexp(n, lambda)
X_2 <- rexp(n, lambda)

curve(dexp(x, rate = lambda), from=0, to=10, col='blue')
curve(dexp(x, rate = lambda), from=0, to=10, col='red', add = TRUE)

Z <- X_1 - X_2

hist(Z, breaks = 200)
```

```{r}
n <- 100000
lambda <- 1
X_1 <- rexp(n, 3*lambda)
X_2 <- rexp(n, lambda)

curve(dexp(x, rate = 3*lambda), from=0, to=10, col='blue')
curve(dexp(x, rate = lambda), from=0, to=10, col='red', add = TRUE)

Z <- X_1 - X_2

hist(Z, breaks = 200)
```

## 3. Let $X$ be a continuous random variable with mean $\mu = 10$ and variance $\sigma^2 = \frac{100}{3}$.  Using Chaebyshev's Inequality, find an upper bound for the following probabilities:

\[Chebyshev's \hspace{2mm} Inequality\]
\[P (|X - \mu| \ge \epsilon) \le \frac{V(X)}{\epsilon^2}\]

```{r}
variance <- 100/3

cheb <- function(epsilon, var = variance) {
  upperBound <- var / epsilon^2
  if (upperBound > 1) {
    upperBound <- 1
  }
  
  return(upperBound)
}
```


>(a) $P(|X - \mu| \ge 2)$

```{r}
cheb(2)
```

>(b) $P(|X - \mu| \ge 5)$

```{r}
cheb(5)
```

>(c) $P(|X - \mu| \ge 9)$

```{r}
cheb(9)
```

>(d) $P(|X - \mu| \ge 20)$

```{r}
cheb(20)
```

