---
title: "Week 7"
author: "Sam Reeves"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Let $X_1, X_2, ..., X_n$ be $n$ mutually independent random variables, each of which is uniformly distributed on the integers from 1 to $k$.  Let $Y$ denote the minimum of $X_i$s.  Find the distribution of $Y$.

This pertains to the first section of Chapter 7 in our probability text: Sums of Discrete Random Variables. Because the variables are distributed on the set of integers, we know this is not a continuous distribution.

I think the following statement provides a hint:

>We shall find it convenient to assume here that these distribution functions are defined for all integers, by defining them to be 0 where they are not otherwise defined.

So for $k \in \{1, 2, ..., k\}$, $P(X = i) = \frac{1}{k}$.

$P(X > 0) = 1$, $P(X > 1) = 1 - \frac{1}{k}, …, P(X > k-1) = \frac{1}{k}, P(X > k) = 0$

The text gives:

\[Z = X+Y\]

\[X = k, Z =z\]
\[Y =  z-k\]

\[P(Z=z) = \sum_{k = -\infty}^{\infty} P(X = k) \cdot P(Y = z-k)\]

and follows with:

\[m_3(j) = \sum_k m_1(k) \cdot m_2(j-k)\]

suggesting that $m_3(x)$ is the distribution function for the random variable $Z = X + Y$.  It is explained that convolutions are commutative and associative... So, we can manipulate the order of these terms.

If $S_n = X_1 + X_2 + ... + X_n$, then $S_n = S_n-1 + X_n$.

So, the distribution of Y is:

\[P(X_i = Y) = \sum_{i=1}^k\frac{(k)^n - (k-i)^n}{k^n}\]

That is, the sum of the ways we can get the minimum $Y$ over the number of possible values of $X_i$.


## 2. Your organization owns a copier.  This machine has a manufacturer's expected lifetime of 10 years. This means that we expect one failure every ten years.  For each of the following: include the probability that the machine will fail after 8 years, the expected value, and the standard deviation.

```{r}
p <- 1/10
n <- 8
```

### a. Model as geometric.  (Hint: the probability is equivalent to not failing during the first 8 years...)

In this week's lecture, we learned that the pgeom function is very precise in that it calculates $p(n) = pq^n$.  These are 8 years of not failing followed by a year of failure.  We can specify lower.tail  = TRUE to indicate $P(X \le )$ as opposed to $P(X > x)$, as given in the documentation for pgeom.

For our purposes, we want $P(X > n) = (1 - p)^{(n + 1)}$, demonstrated:

```{r}
(1-p) ^ (n+1)
pgeom(n, p, lower.tail = FALSE)
```

Borrowing from my work in week 5:

\[E(x) = \sum_{x \in \Omega} xm(x) = \frac{1}{p}\]
\[= 10\]

Standard deviation is $\sigma = \sqrt{\frac{1-p}{p^2}}$

```{r}
sqrt( (1-p) / (p^2) )
```

### b. Model as an exponential.

The probability that the machine will fail after 8 years is:

\[P(X) = 1 - (1 - e^{-pn})\]
\[= e^{-pn}\]

```{r}
exp(1)^(-p * n)
pexp(8, p, lower.tail = FALSE)
```

The expected failure rate is given by the manufacturer to be 10 years.  We can still call this $\frac{1}{p}$.

For $p = \lambda$:

\[\mu_X = \frac{1}{\lambda}\]
\[\sigma^2_X = \frac{1}{\lambda^2}\]
\[\mu_X = \sigma_X\ = \frac{1}{\lambda}\]
\[\sigma_X = 10\]

The standard deviation is equal to the mean.

### c. Model as a binomial.  (Hint: 0 successes in 8 years...)

Borrowing the concepts from Michael Ippolito's recitation in this week's lecture:

\[b(n, p, k) = {n \choose k} p^k (1-p)^{n-k}\]
\[= {8 \choose 0} \frac{1}{10}^0 \frac{9}{10}^8\]
\[= {8 \choose 0} \frac{9}{10}^8\]

According to the documentation for the binomial distribution functions in R, we can use the dbinom function to return the probability density of the function described above...

```{r}
k <- 0

choose(8,0) * (1-p)^8
dbinom(k, n, p)
```

In this case, the expected value is $E(X) = np = 0.8$

The standard deviation for a binomial distribution is:
\[\sigma = \sqrt{np(1-p)}\]

```{r}
sqrt(n * p * (1-p))
```

### d. Model as a Poisson.

According to the documentation for the Poisson distribution functions in R:

\[P(X) = \lambda^x e \frac{-\lambda}{x!}\]

In our case, $\lambda = p$.  So, we can use the ppois function with $x = 0$ to return the probability that there is no failure in a year, then multiply that times itself 8 times to find the probability that there are no failures for the first 8 years.

\[P(0) = \lambda^0 e\frac{-\lambda}{x!}\]
\[0! = 1\]
\[\therefore\]
\[P(0,1,...7) = (e \times -\lambda)^8\]

```{r}
exp(-p)^8
ppois(0, p)^8
```

The expected value is $E(X) = \lambda n$:

```{r}
n*p
```

In the Poisson distribution, the variance is equal to the expected value for one trial, so the standard deviation is equal to the squareroot of the expected value:

\[E(X) = \sigma^2\]
\[\sigma = \sqrt{\lambda n}\]

```{r}
sqrt(n*p)
```

