---
title: "discussion15"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r}
f1 <- function(x,y) {1 / sqrt(y - x^2)}

x <- seq(0, 10, length = 10)
y <- seq(1, 110, length = 10)
z <- outer(x, y, f1)

persp(x, y, z, main = "Some Section of 1/y-x^2", 
      theta = 300, phi = 20, col = "blue", shade = 3)
```

```{r}
f2 <- function(x,y) {cos(y) * sin(x) / x}

x <- seq(-10, 10, length = 100)
y <- seq(-10, 10, length = 100)
z <- outer(x, y, f2)

persp(x, y, z, main = "Some Section of cosy*sinx/x", 
      theta = 320, phi = 30, col = "red", shade = 0.5)
```

12.2 #20

A limit is given. Evaluate the limit along the paths given, then state why these results show the given limit does not exist.

\[\lim_{(x,y) \to (\pi,\pi / 2)} \frac{\sin x}{\cos y}\]

(a) Along $x = \pi$

First, substitute:
\[\lim_{y \to \pi / 2} \frac{\sin pi}{\cos y}\]

Now apply Bernoulli's Rule:

```{r}
D(sin(pi)/cos(y), 'y')
```


(b) Along $y = x - \pi / 2$

\[\lim_{x \to x - pi} \frac{\sin x}{\cos x - \pi}\]

```{r}
D(sin(x)/cos(x-pi), 'x')
```

Evaluating this function for the pair (0,0) gives us negative infinity.... 

```{r}
cos(0) / -sin(0)
```

This shows that $y = 0$ is outside of the domain of the function.  Dividing by zero is not okay in this case.

