---
title: "Multivariable Calculus"
author: "Sam Reeves"
date: "12/10/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mosaicCalc)
```

### 1. Find the equation of the regression line for the given points.  Round any final values to the nearest hundredth, if necessary.

```{r}
x <- c(5.6, 6.3, 7, 7.7, 8.4)
y <- c(8.8, 12.4, 14.8, 18.2, 20.8)

points <- cbind(x,y)

lm(y ~ x, data.frame(points))
```

\[y = 4.26x - 14.8\]

### 2. Find all local maxima, local minima, and saddle points for the function given below.  Write your answer(s) in the form (x, y, z).  Separate multiple points with a comma.

```{r}
f2 <- function(x,y) {24*x - 6*x*y^2 - 8*y^3}

x <- y <- (seq(-10, 10, length = 50))
z <- outer(x, y, f2)

persp(x,y,z, theta = 100, phi = 10, col = 'green', ticktype = 'detailed')
```
We take partial derivatives:

\[f_x(x,y) = -6y^2 + 24\]
\[f_y(x,y) = -24y^2 - 12xy\]

We set the first function to 0, yeilding $y = \pm 2$.  Substituting both values into the second partial derivative yeilds $x = 4$ when $y = -2$ and $x = -4$ when $y = 2$.

So we have two VIPs: $(4,-2,64)$ and $(-4,2,-64)$.  It's unclear (analytically) what type of points they are.  Visually, they appear to be saddle points.


### 3. A grocery store sells two brands of a product, the 'house' brand name and a 'name' brand.  The manager estimates that if she sells the 'house' brand for $x$ dollars and the 'name' brand for $y$ dollars, she will be able to sell $81 - 21x + 17y$ units of the 'house' brand and $40 + 11x - 23y$ units of the 'name' brand.

```{r}
house <- function(x,y) {81 - 21*x + 17*y}
name <- function(x,y) {40 + 11*x - 23*y}

x <- y <- seq(0, 50, length = 50)

z_h <- outer(x,y,house)
z_n <- outer(x,y,name)

persp(x,y,z_h, theta = 130, phi = 10, col = 'red')
persp(x,y,z_n, theta = 130, phi = 10, col = 'blue')
```

> Step 1. Find the revenue function $R(x,y)$.

Let's add the functions together:

\[R(x,y) = R(x) + R(y)\]
\[x(81 - 21x + 17y) + y(40 + 11x - 23y)\]
\[R(x,y) = -21x^2 - 23y^2 + 28xy + 81x + 40y\]

```{r}
R <- function(x,y) {-21*x^2 - 23*y^2 + 28*x*y + 81*x + 40*y}

z_R <- outer(x,y,R)
persp(x,y,z_R, col = 'purple', ticktype = 'detailed')
```


> Step 2. What is the revenue if she sells the 'house' brand for $2.30 and the 'name' brand for $4.10?

```{r}
R(2.3, 4.1)
```


### 4. A company has a plant in Los Angeles and a plant in Denver.  The firm is committed to produce a total of 96 units of a product each week.  The total weekly cost is given by $C(x,y) = \frac{1}{6}x^2 + \frac{1}{6}y^2 + 7x + 25y + 700$, where $x$ is the number of units produced in Los Angeles and $y$ is the number of units produced in Denver.  How many units should be produced in each plant to minimize the total weekly cost?

We start with two equations:
\[f(x,y) = \frac{1}{6}x^2 + \frac{1}{6}y^2 + 7x + 25y + 700\]
\[x + y = 96\]

Substitute the simple one into the long one:

\[x = -y + 96\]
\[f(x,y) = \frac{y^2}{3} - 14y + 2908\]

We find the minimum by taking the derivative of this function and setting it equal to zero.

\[f'(y) = \frac{2y}{3} - 14 = 0\]
\[y = 21\]

75 units should come from LA, 21 from Denver.


### 5. Evaluate the double integral on the given region:

\[\int \int e^{8x + 3y} dA\]
\[R: 2 \le x \le 4 \text{and} 2 \le y \le 4\]

We can first evaluate the inner function on the interval:

\[\int_2^4 \int_2^4 e^{8x + 3y} dy dx\]
\[\int_2^4 \frac{1}{8} e^{3y + 32} - e^{3y +16} dx\]

Then, evaluate the outer function on the same interval:

\[\frac{e^{44} - e^{38} - e^{28} + e^{22}}{24}\]



