---
title: "Homework 3  -- Sam Reeves"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# \[Problem Set 1\]

### 1. What is the rank of the matrix A?

```{r}
A <- matrix(c( 1, 2, 3, 4,
              -1, 0, 1, 3,
               0, 1,-2, 1,
               5, 4,-2,-3), ncol = 4, byrow = TRUE)

pracma::rref(A)
```

The matrix A reduces down to the 4 dimensional Identity Matrix.  Each column is a member of the set of dependent variables D.  So, $r(A) = 4$.  A has "full rank".

### 2. Given an mxn matrix where m > n, what can be the maximum rank?  The minimum rank, assuming the matrix is non-zero?

If a matrix has nullity 0; it has no free variables, its rank is n, as in problem 1.  Imagining another scenario, in which the set D has no members, the rank will be 0.

### 3. What is the rank of matrix B?

```{r}
B <- matrix(c(1, 2, 1,
              3, 6, 3,
              2, 4, 2), ncol = 3, byrow = TRUE)

pracma::rref(B)
```

$D = \{1, 3\}$ \hspace{1cm} $F = \{2\}$

$r(A) = 2$


# \[Problem Set 2\]

### Compute the eigenvalues and eigenvectors of the matrix A.  You'll need to show your work.  You'll need to write out the characteristic polynomial and show your solution.

```{r}
A = matrix(c(1, 2, 3,
             0, 4, 5,
             0, 0, 6), ncol = 3, byrow = TRUE)

pracma::rref(A)
```
The matrix A reduces to the 3 dimensional identity matrix $I_3$.

To find the eigenvalues $\lambda$ of A, we must use the following formula:

\[\det(\lambda I_3 - A) = 0\]
\[or\]
\[\left|\begin{array}{ccc}
    \lambda-1 & 0 & 0 \\
    0 & \lambda-1 & 0 \\
    0 & 0 & \lambda-1
    \end{array}
    \right| = 0\]

So, using the $Rule of Sarrus$, we can arrive at the characteristic polynomial by repeating the first $n-1$ columns of A in reduced row-echelon form in columns $n$ through $2n-1$.  Then we can add the sum of the three full length diagonal products, and substract the sum of the three opposite diagonal products.

Performing this action, we get:

\[((\lambda-1)(\lambda-1)(\lambda-1) + 0 + 0) - (0 + 0 + 0) = 0\]
\[(\lambda-1)^3 = 0\]
\[\lambda = 1\]

To compute the eigenspace, the set of vectors that satisfies $(\lambda I_n - A)\vec{v} = \vec{0}$, we can use the following:

\[E(A)_\lambda = N(\lambda I_n - A)\]
\[E(A)_\lambda = N(\left[\begin{array}{ccc}
    0 & 0 & 0 \\
    0 & 0 & 0 \\
    0 & 0 & 0
    \end{array}
    \right])\]
\[E(A)_\lambda = \left\{
    \left[\begin{array}{c}
    v_1 \\
    v_2 \\
    v_3
    \end{array}\right] = t
    \left[\begin{array}{c}
    0 \\
    0 \\
    0
    \end{array}\right],
    t \in \mathbb{R} \right\}\]
