---
title: "notes"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

There are two main libraries for multinomial models in R:

- foreign
- nnet

There are other mechanisms for tuning a multinomial model, but nnets are quite nice.

```{r, eval=FALSE}
sub <- mydata[]
temp <- matrix(rep(0, 28*28*10), 28, 28)

for (i in 1:10){
  temp[i,] <- matrix(as.numeric(mydata[i, 2:ncol(mydata)]), 28, 28)
}
```

```{r}
par(mfrow = c(2, 2)) # set up graphs in a 2x2 matrix
s <- seq(-10, 10, by = 0.01)

fx <- function(x) {3 * (x+1) * (x-4) * (x+5)}
a <- fx(s)
plot(a ~ s, type='l')

fy <- function(y) {y ^ 2}
a <- fy(s)
plot(a ~ s, type='l')

fz <- function(z) {log(z)}
a <- fz(s)
s <- seq(0, 10, by = 0.005)
plot(a ~ s, type='l')

```

Does not pass the line test, isn't a function:
```{r, eval=FALSE}
fqplus <- function(qplus) {sqrt(qplus)}
fqneg <- function(qneg) {-sqrt(qneg)}
s <- seq(0, 10, by=0.01)
a <- fqplus(s)
b <- fqneg(s)
s2 <- c(s,s)
plot(figukrjebcfefkctkunjrftculn
     n ~ s2, type='p')
```


## Evaluating limits numerically in R

```{r}
library(rootSolve)
fx <- function(x) {3 * (x+1) * (x-4) * (x+5)}
s <- seq(-10, 10, by=0.01)
a <- fx(s)
plot(a ~ s, type='l')

yint <- fx(0)
xint <- uniroot(fx, c(-100, 100))
points(xint$root, yint, col='red', pch=21)

xint$root
yint
```

### Derivatives

We're going to just take super small steps to see what the change is.  This is similar to Archimedes process of exhaustion... Change in y over change in x, nuff said.

\[\frac{dy}{dx} = \frac{d}{dx} (f(x)) = \frac{df}{dx}\]

We can use library(Deriv):

```{r}
# symbolic derivative of loc(x)
library(Deriv)

myf1 <- function(x)sin(x)
Deriv(myf1)
```


```{r}
# symbolic derivative of 1/x
myf2 <- function(x)1/x
ans <- Deriv(myf2)

ans
ans(3) # solution evaluated at x=3
```

```{r}
# Power Rule,  px^(p-1)
myf1 <- function(x)x**n
Deriv(myf1)

# Addition Rule, dx(f) + dx(g)
myf2 <- function(x)3 * x + (3 * x^2 + 1)
Deriv(myf2)

# Expponential Function, Identity
myf3 <- function(x)exp(x)

# Logarithm Function, 1/x
myf4 <- function(x)log(x)
Deriv(myf4)
```

```{r}
# Product Rule, f'g + g'f
myf5 <- function(x)3 * x * (3 * (x^2) + 1)
Deriv(myf5)

# Quotient Rule, f/g, (f'g - g'f)/g^2
myf6 <- function(x)3 * x / (3 * (x^2) + 1)
Deriv(myf6)

# Chain Rule, f'u * g'x
myf7 <- function(x)3 * (x^2 + 2)^2
Deriv(myf7)
```


### Max and Min of Functions

```{r, message=FALSE}
# f(x) = x^3 - 3*x^2 - 9*x + 5   for  -2 <= x <= 4

library(mosaic)
library(mosaicCalc)
library(Deriv)
library(rootSolve)

fx <- function(x) {x^3 - 3*x^2 - 9*x + 5}
s <- seq(-2, 4, by=0.1)
a <- fx(s)

data <- cbind(s, a)
plot(data, type='l',
     main='Solve for Max between -2 and 4')

min(a)
max(a)

first <- Deriv(fx)
roots <- uniroot.all(first, c(-2, 4))
roots
```

### Integrals are anti-derivatives!

\[\int_a^b f(x) dx = \lim_{n \to \infty} \sum_{i=1}^{n} \Delta x\]

```{r}
# R base function integrate

fx <- function(x)-2
integrate(Vectorize(fx), 1, 4) #some functions puke without vectorize

fx <- function(x)exp(x^2 + x)
integrate(fx, 1, 2)
```

### Fundamental Theorem of Calculus

Part 1:

\[\int_a^b F'(x)dx = F(b) - F(a)\]

Part 2:

\[\text{If } A(x) = \int_a^x f(t)dt\]
\[A'(x) = \frac{d}{dx} \int_a^x f(t)dt = f(x)\]


```{r}
# Constant Multiple Rule, int(k*f) = k* int(f) + C
antiD(5 ~ x)

# Sum Rule (separable)
antiD(x-2 + x^2 ~ x)

# Power Rule, x^n -> (x^(n+1)) / (n+1)
antiD(x^3 ~ x)

# exp(x) => exp(x) + C
antiD(exp(x) ~ x)

# 1/x => ln|x| + C
antiD(1/x ~ x)
```

#### Substitution

\[u = f(x)\]
\[x = f^-1 (u)\]
\[du = f'(x)dx\]
\[dx = \frac{du}{f'(x)}\]

#### By parts
\[\int (udv) = uv - \int (vdu)\]


### Two (and more) Variable Problems

Use the cubature library, it will do an adaptive integration for definite integrals.

```{r}
library(cubature)

fxyz <- function(x) {(x[1]-2)^2 + (x[2]+3)^2 + (x[3]-4)^2 - 9}

adaptIntegrate(fxyz,
               lower = c(-10, -10, -10),
               upper = c(10, 10, 10))
```






