---
title: "Part 1"
author: "Sam Reeves"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(igraph)
library(matrixcalc)
```


# Part 1. Playing with Pagerank

## 1. The A and B Matrices (5)

### Form the $A$ matrix.  Then, introduce decay and form the $B$ matrix as we did in the course notes.

```{r}
n = 6

A <- matrix(c(0, 1/2, 1/2, 0, 0, 0,
              0, 0, 0, 0, 0, 0,
              1/3, 1/3, 0, 0, 1/3, 0,
              0, 0, 0, 0, 1/2, 1/2,
              0, 0, 0, 1/2, 0, 1/2,
              0, 0, 0, 1, 0, 0),
            ncol = n, byrow = FALSE)
```

\[B = 0.85 \times A + \frac{0.15}{n}\]

This is the formula given in the text, but it doesn't produce the values/effects we are looking for because the second row has all zero entries.  This is not a regular markov chain, and the values will converge only to zero.

This is a normal situation, however, and in the text, the suggested solution to this problem is to have an even probability of randomly jumping to any page even though there are no links out from the second page.  We see that all the columns add up to a total probability of 1 except for column number 2:

```{r}
B <- 0.85 * A + 0.15 / n
colSums(B)
```

So, because of this sad situation, our values will not converge to anything but zero.  The matrix will also not have an eigenvalue of 1.  We have to alter the original matrix, such that the 6 page link probabilities from page 2 are even and add up to 1.

```{r}
A[,2] <- rep(1/6, 6)
B <- 0.85 * A + 0.15 / n

colSums(A)
```


## 2. The Rank Vector (5)

### Start with a uniform rank vector $r$ and perform power iterations on $B$ till convergence.  That is, compute the solution $r = B^n \times r$.  Attempt this for a sufficiently large $n$ so that $r$ actually converges.

```{r}
r <- rep(1/n, n)
matrix.power(B, 1000) %*% r
```


## 3. Eigen-Decomposition (10)

### Compute the eigen-decomposition of $B$ and verify that you indeed get an eigenvalue of 1 as the largest eigenvalue and that its corresponding eigenvector is the same vector that you obtained in the previous power iteration method.  Further, this eigenvector has all positive entries and it sums to 1.

```{r}
eig_B <- eigen(B)

eig_B$values[1]
eig_B$vectors[,1]/2
```

I do not understand why each value in the eigenvector is excatly double the converged values above...  This seems to work perfectly, but for the wrong reasons.

## 4. Graph Algorithm (10)

### Use the $graph$ package in R and its $page.rank$ method to compute the PageRank of the graph as given in $A$.  Note that you don't need to apply decay.  The package starts with a connected graph and applies decay internally.  Verify that you do get the same PageRank vector as the two approaches above.

```{r}
edges <- matrix(c(1, 2, 1/2,
                  1, 3, 1/2,
                  3, 1, 1/3,
                  3, 2, 1/3,
                  3, 5, 1/3,
                  4, 5, 1/2,
                  4, 6, 1/2,
                  5, 4, 1/2,
                  5, 6, 1/2,
                  6, 4, 1),
                ncol = 3, byrow = TRUE)

edges_1 <- c(1,2,1,3,3,1,3,2,3,5,4,5,4,6,5,4,5,6,6,4)

g <- graph(edges_1) 
g <- set_edge_attr(g, "w", value = edges[,3])

plot(g)
page.rank(g)
```

