---
title: "Taylor Series Discussion"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pracma)
library(mosaic)
library(mosaicCalc)
```

## Ch. 8.8 #31

Approximate the value of the given definite integral by using the ﬁrst 4 nonzero terms of the integrand’s Taylor series.

\[\int_0^{\sqrt{\pi}} \text{sin} (x^2)dx\]

 ---
 
 First, we find the taylor series...

\[\text{sin} x = \sum_{n=0}^{\infty} (-1)^n \frac{x^{2n+1}}{(2n+1)!}\]

\[x - \frac{x^3}{3!} + \frac{x^5}{5!} - \frac{x^7}{7!}\]
 
\[\text{sin} (x^2) = \sum_{n=0}^{\infty} (-1)^n \frac{x^{4n+2}}{(2n+1)!}\]

\[x^2 - \frac{x^6}{3!} + \frac{x^{10}}{5!} - \frac{x^{14}}{7!}\]

Then we define the function and integrate each term.

```{r}
antiD(x^2 ~ x)
antiD(-x^6 / fact(3) ~ x)
antiD(x^10 / fact(5) ~ x)
antiD(-x^14 / fact(7) ~ x)
```

\[C + \frac{x^3}{3} + \frac{x^7}{42} + \frac{x^{11}}{1320} + \frac{x^{15}}{75600}\]

Then, we evaluate it for the interval:

```{r}
f <- function(x) {x^3/3 - x^7/42 + x^11/1320 - x^15/75600}

f(sqrt(pi)) - f(0)
```


```{r}
f <- function(x) x - (x ** (3/2) / factorial(3)) + (x ** (5/2) / factorial(5)) - (x ** (7/2) / factorial(7))

f((pi^2)/4)

f <- function(x) 1 - sqrt(x)^2/fact(2) + sqrt(x)^4/fact(4) - sqrt(x)^6/fact(6)

integrate(f, 0, pi^2/4)
```

