---
title: "MIT Taylor Series notes"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Taylor Series
Stacking blocks.... start from the top:

The block is length 2, so that the leftmost part of the second block is its midpoint, 1.  The third block is now at 3/2.

```{r}
series1 <- c(0, 1, 3/2)
```

$C_n$ is the center of mass of the block tower... For N blocks, the new block $N + 1$ has a center at $C_n +1$.  Now, we can calculate the center of mass (for just the x coordinate).

\[C_N = \frac{NC_N + C_N + 1}{N +1} = \frac{(N + 1)C_N + 1}{N + 1}\]
\[C_{N+1} = C_N + \frac{1}{n+a}\]

\[C_1 = 1\]
\[C_2 =  1 + \frac{1}{2}\]
\[C_3 = 1 + \frac{1}{2} + \frac{1}{3}\]

\[C_N = 1 + \frac{1}{2} + \frac{1}{3} + \frac{1}{4} + ... + \frac{1}{N}\]

\[C_N = S_N\]
\[ln N < S_N < (ln N) + 1\]

As $N \to \inf$, $ln N \to \inf$, and $S_N \to \inf$.

---

We can make a tower with an arbitrary width, we can determine exactly how many blocks are needed.

\[C_{N+1} = \frac{NC_N + 1(C_N +1)}{N + 1} = \frac{(N+1)C_N + 1}{N + 1}\]
\[C_{N+1} = C_N + \frac{1}{N+1}\]

\[C_1 = 1\]
\[C_2 = 1 + \frac{1}{2}\]
\[C_3 = C_2 + \frac{1}{3} = 1 + \frac{1}{2} + \frac{1}{3}\]
\[C_N =  1 + \frac{1}{2} + \frac{1}{4} + \frac{1}{4} + ... + \frac{1}{N}\]

\[C_N = S_N\]
\[ln N < S_N < (ln N) + 1\]

As $N \to \infty$, $ln N \to \infty$ and $S_N \to \infty$.

\[N = e^24\]
\[(3cm) e^24 \approx 8 \times 10^8 m\]
\[\text{twice the distance to the moon}\]

 ---
 

### Power Series

\[1 + x + x^2 + x^3 + ... = \frac{1}{1-x}\]

suppose:

\[1 + x + x^2 + ... = S\]

multiply by x:

\[x + x^2 + x^3 + ... = Sx\]

subtract one from the other:

\[1 = S - Sx = S(1-x)\]
\[\frac{1}{1-x} = S\]

This reasoning is basically correct, but is incomplete because it requires that S exists...

e.g.
\[x =1, 1 + 1 + 1 + ... = S\]
\[1 + 1 + 1 = S \times 1\]
\[\infty - \infty = \infty - \infty\]

 ---

### General Power Series

\[ a_0 + a_1 x + a_2 x^2 + a_3 x^3 + ...\]
\[= \sum_{n=0}^{\infty} a_n x^n\]

\[|x| < R \text{radius of convergence}\]
\[-R < x < R\]

where the series converges...

\[|x| > R, \sum a_n x^n \text{diverges}\]

(and $|x| = R very delicate)

What does matter:

$|a_n x^n| \to 0$ exponentially fast for $|x| < R$

\[|a_n x^n| \nrightarrow 0\] for $|x| > R$

Series are flexible enough to represent all the functions we know...
In this form, they are computationally available.

\[1 + x + x^2 + x^3 + ... = \frac{1}{1-x}\]

#### Rules for convergent power series:

Just like polynomials.

\[f(x) + g(x), f(x)g(x),\]
\[f(g(x)), f(x)/g(x)\]

the last two are interesting because we differentiate and integrate them:

\[\frac{d}{dx} f(x), \int f(x) dx\]


 ---
 
 \[\frac{d}{dx}(a_0 + a_1 x + a_2 x^2 + a_3 x^3 + ...) = a1 + 2a_2 x + 3a_3 x^2 + ...\]
 
 \[\int (a_0 + a_1 x + a_2 x^2 + ...)dx = c + a_0 x + a_1 \frac{x^2}{2} + a_2 \frac{x^3}{3} + ...\]


### Taylor's Formula

\[f(x) = \sum \frac{f^{(n)} (0)}{n!} x^n\]

\[f'(x) = a_1 + 2 a_2 x + 3 a_3 x^2 + ...\]
\[f''(x) = 2 a_2 + 3 \cdot{2 a_3 x} + ...\]
\[f'''(x) = 3 \cdot{2 a_3} + 4 \cdot{3} \cdot{2 a_4x} + ...\]

\[f'''(0) = 3 \cdot{2 a_3}\]
\[\frac{f'''(0)}{3 \cdot{2} \cdot{1}} = a_3\]

In general:

\[a_n = \frac{f^{(n)} (0)}{n!}\]

\[n! = n \cdot{(n-1)} \cdot{(n-2)} ... 1\]
\[0! = 1\]

\[f(x) = e^x, f'(x) = e^x, f''(x) = e^x\]
\[...\]
\[\left. f^{(n)}(x) = e^x \right|_{x=0} = 1\]

\[e^x = \sum_{n=0}^{\infty} \frac{1}{n!} x^n\]

\[e = e^1 = 1 + 1 + \frac{1}{2!} + \frac{1}{3!} + \frac{1}{4!} + ...\]

\[sinx \approx x\]
\[cosx \approx 1 - \frac{x^2}{2}\]

\[sinx = x - \frac{x^3}{3!} + \frac{x^5}{5!} - \frac{x^7}{7!} + ...\]
\[cosx = 1 - \frac{x^2}{2!} + \frac{x^4}{4!} - \frac{x^6}{6!} + ...\]