---
title: "Power Plant Capacity"
author: "Sam Reeves"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here we have a dataset containing info concerning about 30k power plants globally.  We are going to try to use all the variables given to predict the capacity in megawatts of a powerplant on Earth. That is, we will make a generalized linear model that describes the dependent variable capacity.

```{r, message=FALSE}
library(dplyr)
library(geosphere)

f <- read.csv("power.csv") %>%
  tibble() %>%
  select(c(Country, Capacity..MW., Latitude, Longitude, Primary.Fuel, Owner, Source)) %>%
  rename(Cap = Capacity..MW.,
         Fuel = Primary.Fuel,
         Lat = Latitude,
         Long = Longitude) %>%
  arrange(Lat)

head(f)
```

### Dichotomous term

Since there is no dichotomous term given automatically by the dataset, I suppose we will have to make one out of latitude.  There are substantially more power plants in the Northern hemisphere than the southern, and I believe that adding this feature should fall in line with other global economic studies.  Frequently, development economists discuss trade and infrastructure in terms of the "global north" or "global south".

```{r, fig.height=3, fig.width=5}
f <- mutate(f, North = ifelse(Lat > 0, 1, 0))
hist(f$Lat)
```

\clearpage

### Quadratic term

There isn't an obvious choice for a quadratic term either.  So, I want to combine this dataset with another that contains information on many of the world's cities.  We will relate each power plant with the nearest city with a population over 1M.

To compute the distance, we can use a number of functions from the geosphere package.  I have chosen the Vincenty Ellipsoid function, whose defaults are set to the shape of the Earth.  These results are very accurate

```{r}
g <- read.csv("worldcities.csv") %>%
  tibble() %>%
  filter(population > 1000000) %>%
  rename(pop = population,
         type = capital) %>%
  select(c(city, lat, lng, pop, type)) %>%
  arrange(lat)

head(g)

findNear <- function(f, g) {
  nearest <- tibble()
  for (i in 1:nrow(f)) {
    cities <- tibble()
    p1 <- c(f[i,]$Long, f[i,]$Lat)

    for (j in 1:nrow(g)) {
      p2 <- c(g[j,]$lng, g[j,]$lat)
      dist <- distVincentyEllipsoid(p1, p2)
      cities <- rbind(cities, cbind(g[j,], dist))
    }
    nearest <- rbind(nearest, cities[which.min(cities$dist),])
  }
  return(nearest)
}
```

Because this process is so computationally intensive, we have have to sample the data...

```{r}

```

### Dichotomous vs. quantitative interaction term

For this term, we will use the binary term North and the population of the nearest large city.

```{r}
m <- cbind(f, h) %>%
  mutate(dist2 = dist^2,
         dvq = North * dist)
```

### Interpret all coefficients

```{r}
lm1 <- lm(Cap ~ North + Fuel + Owner + Source + pop + type + dist + dist2 + dvq, m)
```

### Conduct residual analysis

### Was the linear model appropriate? Why or why not?