---
output:
  pdf_document: default
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(forecast)
library(ggplot2)
lynx <- datasets::lynx

lynx.model <- lm(lynx ~ c(1821:1934), lynx)
which(lynx>5000) + 1821
```

```{r}
plot(lynx)
```

There is a cycle of 38 year peaks.  These peaks are about twice the size of the relative peaks between, which happen every 9.5 years.

```{r}
plot(lynx.model$residuals)
```

It is interesting that the peak of 1905 allowed for two consecutive years of 5000+ trappings.  It's definietly cyclical, but is there a trend?

```{r}
lynx.trend <- na.omit(ma(lynx, order = 38))
```

In this example, taking outliers away would be an absolute sin.

```{r}
plot(lynx.trend)
```

```{r}
lynx.short <- data.frame(cbind(lynx[20:95], c(1840:1915)))
colnames(lynx.short) <- c("trappings", "year")

trend.model <- lm(lynx.trend ~ c(1840:1915), lynx.trend)

lynx.adj <- (lynx.trend * trend.model$coefficients[2]) + trend.model$coefficients[1]

lynx.adj <- lynx.adj - min(lynx.adj)
```

```{r, messaege=FALSE, warning=FALSE, echo=FALSE}
ggplot(lynx.short, aes(x = year,
                       y = lynx.short$trappings)) +
  geom_abline(intercept = trend.model$coefficients[1],
              slope = trend.model$coefficients[2],
              color = "white") +
  geom_point(aes(y = lynx.adj)) +
  geom_point(aes(color = lynx.short$trappings)) +
  theme_dark() +
  theme(legend.position = "none") +
  labs(title = "Lynx trappings and seasonally adjusted trend",
       x = "", y = "")
```

This is not a relationship that can be modeled with a simple linear regression, but it looks like there is a trend line during this period...
